---
title: "How does a pendulum's length affect its time period?"
author: "Heather Wang, Qihan Liu, Runxi Yu, Vincent Bae"
output: html_document
---

## Introduction

![Simple gravity pendulum](Simple_gravity_pendulum.svg)

A pendulum

## Experiment design

## Analysis
 
```{r}
data <- read.csv("data.csv")
data
summary(data)
```

In order to gain a general sense of the data we have obtained, let us plot the
data. Since each independent variable has multiple trials and therefore has
multiple dependent data points, we use boxplots:

```{r}
boxplot(time_10_periods ~ length_cm, data = data,
	main = "Effect of Length on Time",
	xlab = "Length (cm)",
	ylab = "Time (10 periods)")
```

The data is roughly linear. So,

```{r}
model <- lm(time_10_periods ~ length_cm, data = data)
summary(model)
```

We shall run various diagnostic plots of this linear regression model to
confirm whether it is truly a linear regression:

```{r}
plot(model)
```

```{r}
pairwise.t.test(data$time_10_periods, data$length_cm, p.adjust.method = "bonferroni")
```

## Results

## Informal bibliography

I'm not really sure how to get BibLaTeX working with R Markdown yet, so here's
just an itemized list of references without any particular bibliography format.

* Simple Gravity pendulum, by Chetvorno, public domain, https://commons.wikimedia.org/w/index.php?curid=5276335
