---
title: "How does a pendulum's length affect its time period?"
author: "Heather Wang, Qihan Liu, Runxi Yu, Vincent Bae"
output: html_document
---

## Introduction

In this experiment, we investigate whether changing the length of a pendulum's
string causes changes in its period (the time it takes to complete one full
swing back and forth).

![Simple gravity pendulum](Simple_gravity_pendulum.svg)

Theory suggests that for sufficiently small amounts angles, the pendulum
follows $T = 2\pi\sqrt{\frac{L}{g}}$ where $T$ is the period, $L$ is the
length, and $g$ is the acceleration due to gravity. Therefore, the theory
suggests that changes in the string's length do affect the time period.

Therefore, our hypotheses shall be:
* **Null hypothesis:** Changes in the string's length do not affect the time
  period.
* **Alternate hypothesis:** Changes in the string's length do affect the time
  period.

## Treatment

The treatment in this experiment is the pendulum length, for which we suggested
$10\,\mathrm{cm}$, $15\,\mathrm{cm}$, $20\,\mathrm{cm}$, and $25\,\mathrm{cm}$.

Random assignment is not applicable in this context because the experiment is
designed as a deterministic system in a controlled environment; all extraneous
variables, such as the mass of the bulb, the gravitational field strength of
the earth, were held constant as far as possible. Each length was tested in
four trials to reduce measurement errors.

There are some limitations on generalizability. First, although the theoretical
formula assumes small-angle oscillations, the initial release angles in this
experiment were relatively large. This introduces minor discrepancies due to
nonlinear effects at larger angles. Additionally, the results may only
generalize to pendulums with similar construction materials and in similar air
conditions (e.g., negligible air resistance).

## Subjects

The subjects are the pendulums constructed by tying two 1 CNY coins to a length
of string. The same type and number of coins were used throughout the
experiment. A piece of string was suspended from a fixed support, and the coins
were attached to the end of the string to act as the pendulum bob. The pendulum
was released from a fixed angle, and the time it took to complete one period
was recorded using a stopwatch.

2 coins were chosen as they are readily available, of a fixed mass, and provide
enough mass for the pendulum to swing. The subjects are ideal for the study
since the pendulum system is a physical object rather than a social or
biological population. They directly represent the physical system being
investigated and allow for controlled manipulation of the treatment variable
while ensuring other factors are constant. There might be limitations in
generalizing the results to all pendulums. The findings will directly apply to
pendulums made with similar materials and in similar settings, but may not be
generalizable to pendulums with different bobs, materials, or conditions. That
said, while the exact timings and results from this experiment is not
applicable to all pendulums, the trend and relationship between pendulum length
and period are generalizable. 

## Analysis

### Data overview
 
```{r}
data <- read.csv("data.csv")
data
summary(data)
```

Treating the length as discrete groups, let us estimate the mean and standard
deviation of each group.

```{r}
aggregate(time_10_periods ~ length_cm, data = data, function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
```

This allows us to see that there is a significant Difference in Means betwen
groups, of approximately 0.6 between neighbouring groups, where the standard
deviation is roughly around 0.1 to 0.2.

In order to gain a general sense of the data we have obtained, let us plot the
data. A box plot is used, to present the information aggregated by group above.

```{r}
boxplot(time_10_periods ~ length_cm, data = data,
	main = "Effect of Length on Time",
	xlab = "Length (cm)",
	ylab = "Time (10 periods)")
```

### Linear regression

```{r}
model <- lm(time_10_periods ~ length_cm, data = data)
summary(model)
```

We shall run various diagnostic plots of this linear regression model to
confirm whether it is truly a linear regression:

```{r}
plot(model)
```

There are insufficient sample points for the diagnostic plots to be accurate,
but they seem to be acceptable for this sample size. It is however somewhat
unclear what causes the enlarged residues for large theoretical quantities.

Given that a linear regression is an appropriate estimate as suggested by the
diagnostic tests (even though the relationship is theoretically supposed to be
that of a square root), its $p$-value $4.52 \times 10^{-12}$ suggests that it
is extremely unlikely to obtain our result (or more extreme results) given that
the null hypothesis is true, therefore suggesting with high confidence that the
null hypothesis is false and that the alternate hypothesis is true. Therefore:
changes in the string's length do affect the time period.

## T-tests

Another way to confirm the statistical significance of the difference in means
is via Welch's $t$-test, not assuming equal variances.

```{r}
group_10 <- data$time_10_periods[data$length_cm == 10]
group_15 <- data$time_10_periods[data$length_cm == 15]
group_20 <- data$time_10_periods[data$length_cm == 20]
group_25 <- data$time_10_periods[data$length_cm == 25]
```

```{r}
print(t.test(group_10, group_15, var.equal = FALSE))
print(t.test(group_15, group_20, var.equal = FALSE))
print(t.test(group_20, group_25, var.equal = FALSE))
```

Again, the $p$-values are extremely small, suggesting that it is extremely
unlikely to obtain our result (or more extreme results) given that the null
hypothesis is true, therefore suggesting with high confidence that the null
hypothesis is false and that the alternate hypothesis is true. Therefore:
changes in the string's length do affect the time period.

## Estimated effects

Are effects estimated correctly? Are they described correctly? Is their substantive significance discussed, ideally with a relevant benchmark?

## Statistical Significance

Are p-values presented and discussed correctly? A full score would discuss why the p-value is obtained, i.e. what about the study led to the p-values estimated.

*I think this is done*

## Informal bibliography

I'm not really sure how to get BibLaTeX working with R Markdown yet, so here's
just an itemized list of references without any particular bibliography format.

* Simple Gravity pendulum, by Chetvorno, public domain, https://commons.wikimedia.org/w/index.php?curid=5276335
